---
title: "An Analysis of Des Moines Real Estate"
author: "Samuel Burer"
date: "Updated: `r format(Sys.time(), '%B %e, %Y')`"
output: slidy_presentation
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

##Introduction

With travel being on everyone's mind during the summer we wanted to take this time to look at a popular vacation spot and see how we can help travelers save money and time for next year's vacation.  The days are gone when traveling all you have to choose from are resorts and hotels, over the last 5 years the idea of Airbnb's have gotten more attractive.  Today's traveler is more adventurous and money conscience, so we want to help today's traveler by showing them that (if they vacation in Hawaii) they know they are getting what they want, at the price the want, near the amenities they are looking for.   This project will use the data from http://insideairbnb.com/get-the-data.html. This is a public domain site that shows data for Airbnb's from around the world, but we will focus only on the data from Hawaii.   It will help us determine and pin down the best area for a traveler to go based on cost, size of the room rented, and nearby amenities.  Our goal is to answer three questions that will help the average traveler make an educated decision on where to stay.   
1. How does price vary with location? We believe that the closer the Airbnb is to the beach, the more expensive the stay will be. 
2. How does the rating and cost of the Airbnb effect it's success?  We can surmise that a better rating will prove more successful than a lower rating despite of the location. 
Lastly, we will answer the question of the best Airbnb to rent based on how long it has been operational.  With these questions answered we feel we will be able to successfully determine the top five Airbnb's in Hawaii.


```{r, echo=FALSE,  comment=NULL, warning=FALSE}
# Load packages
suppressPackageStartupMessages(library(readr))                     
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(choroplethr))
suppressPackageStartupMessages(library(choroplethrMaps))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
```

## Hawaiian Islands

```{r, echo=FALSE,  comment=NULL, warning=FALSE, message=FALSE}

# Load population data by county
data(df_pop_county)

# Store the data in a shorter name
dpc <- df_pop_county


# Create, store, and view plot
p <- county_choropleth(dpc, state_zoom = "hawaii")
p
```

```{r, echo=FALSE,  comment=NULL, warning=FALSE, message=FALSE}
df <- read_csv("listings.csv")
df4 <- read_csv("listings.csv.gz")

#Modify Columns
my_col_types = cols(
  id = col_integer(),
  name = col_character(),
  host_id = col_integer(),
  host_name = col_character(),
  neighbourhood_group = col_character(),
  neighbourhood = col_character(),
  latitude = col_double(),
  longitude = col_double(),
  room_type = col_character(),
  price = col_integer(),
  minimum_nights = col_integer(),
  number_of_reviews = col_integer(),
  last_review = col_date(format = ""),
  reviews_per_month = col_double(),
  calculated_host_listings_count = col_integer(),
  availability_365 = col_integer()
)

#Final Read
df <- read_csv("listings.csv", col_types = my_col_types)


```

## How many AirBnb'S are there per Island

```{r, echo=FALSE,  comment=NULL, warning=FALSE, message=FALSE}

####################################################################################################################
#Factor Islands
####################################################################################################################
df$neighbourhood_group <- factor(df$neighbourhood_group)        
####################################################################################################################

####################################################################################################################
#Review Scores
####################################################################################################################
df$reviewscores <- df4$review_scores_rating
####################################################################################################################


####################################################################################################################
#Add a Column for the Age of Ownership
####################################################################################################################
df <- mutate(df, current_year = as.integer(format(Sys.Date(), "%Y")))
df <- mutate(df, hosttime = as.integer(format(df4$host_since, "%Y")))
df <- mutate(df, age = current_year - hosttime)
####################################################################################################################

####################################################################################################################
#Factor the Price and turn it into a number
####################################################################################################################
df$price <- factor(df$price)
df$price <- as.numeric(df$price)
####################################################################################################################

####################################################################################################################
#Add a column for host who are identified as SuperHost
####################################################################################################################
df <- mutate(df, great_host = df4$host_is_superhost)            #Great Hosts per Island
#df <- mutate(df, great_host = factor(as.integer(df4$host_is_superhost == "t")))
#df$great_host
#class(df$great_host)
df$great_host <- factor(df$great_host)
#levels(df$great_host)
levels(df$great_host) <- c("Not a Great Host", "Great Host")

#print(levels(df$great_host))
####################################################################################################################

####################################################################################################################
#Add a column for east and west side of island for Hawaii
####################################################################################################################
df <- mutate(df, sideofisland = ifelse(longitude<(-155.5),'east','west'))
####################################################################################################################

####################################################################################################################
# Remove records with even a single NA
####################################################################################################################
df <- df[complete.cases(df), ]        
####################################################################################################################

table(df$neighbourhood_group)

###################################################################################################################
# Bar Chart: What island has the most AIRBNB
###################################################################################################################
p <- qplot(neighbourhood_group, data = df, geom = "bar", xlab = "Island", ylab = "QTY of BnB",
           fill = I("yellow"),                     # Add a light blue fill
           color = I("red"),                          # Make color of rectangles red
           alpha = I(0.5))                            # Set 50% transparancy
p <- p + ggtitle("QTY of AirBnB per Island")
p <- p + theme(panel.border = element_rect(linetype = "solid", color = "black", fill = NA))
p <- p + theme(panel.border = element_rect(linetype = "solid", color = "black", fill = NA))
p <- p + theme(plot.title = element_text(colour="grey20",size=20,face="bold"))
p <- p + theme(axis.text = element_text(face="bold"))
p <- p + theme(axis.title.y = element_text(colour="grey20",size=20,face="bold"),
               axis.text.x = element_text(colour="grey20",size=20,face="bold"),
               axis.text.y = element_text(colour="grey20",size=20,face="bold"),  
               axis.title.x = element_text(colour="grey20",size=20,face="bold"))
p
###################################################################################################################

```

## How many AirBnb'S are there per Island

```{r, echo=FALSE,  comment=NULL, warning=FALSE, message=FALSE}


####################################################################################################################
# Picture of all Islands: Overall Picture of all AirBnB
####################################################################################################################
p <- qplot(longitude, latitude, data = df, color = neighbourhood_group, na.rm = TRUE, xlab="Longitude", ylab="Latitude") 
p <- p + scale_color_hue(name = "Island")
p <- p + ggtitle("Overview of AirBNB per Island")
p <- p + theme(panel.border = element_rect(linetype = "solid", color = "black", fill = NA))
p <- p + geom_text(aes(label=ifelse(latitude<(18.99),as.character(neighbourhood_group),'')),hjust=1.5,vjust=-4, color = "blue", size= 8)
p <- p + geom_text(aes(label=ifelse(latitude>(21.7) & latitude<(21.8),as.character(neighbourhood_group),'')),hjust=-.5,vjust=1, color = "blue", size = 8)
p <- p + geom_text(aes(label=ifelse(latitude>(21.86) & latitude<(21.87),as.character(neighbourhood_group),'')),hjust=-.2,vjust=1, color = "blue", size = 8)
p <- p + geom_text(aes(label=ifelse(latitude>(21.13) & latitude<(21.14),as.character(neighbourhood_group),'')),hjust=-.5,vjust=4, color = "blue", size = 8)
p <- p + theme(plot.title = element_text(colour="grey20",size=20,face="bold"))
p <- p + theme(axis.text = element_text(face="bold"))
p <- p + theme(axis.title.y = element_text(colour="grey20",size=20,face="bold"),
               axis.text.x = element_text(colour="grey20",size=20,face="bold"),
               axis.text.y = element_text(colour="grey20",size=20,face="bold"),  
               axis.title.x = element_text(colour="grey20",size=20,face="bold"))
p


```

## Years of Ownership per Island

```{r, echo=FALSE,  comment=NULL, warning=FALSE, message=FALSE}

###################################################################################################################
#Bar Chart : Years of Ownership per Island
###################################################################################################################
p <- qplot(age, data = df, geom = "bar", xlab = "Years of Ownership", ylab = "QTY of BnB",
           fill = I("yellow"),                     # Add a light blue fill
           color = I("red"),                          # Make color of rectangles red
           alpha = I(0.5))                            # Set 50% transparancy
p <- p + ggtitle("Overall AVG.Years of Ownership = 3.78 years")
p <- p + theme(panel.border = element_rect(linetype = "solid", color = "black", fill = NA))
p <- p + theme(plot.title = element_text(colour="grey20",size=20,face="bold"))
p <- p + theme(axis.text = element_text(face="bold"))
p <- p + theme(axis.title.y = element_text(colour="grey20",size=20,face="bold"),
               axis.text.x = element_text(colour="grey20",size=20,face="bold"),
               axis.text.y = element_text(colour="grey20",size=20,face="bold"),  
               axis.title.x = element_text(colour="grey20",size=20,face="bold"))
p

```

## Price distribution

```{r, echo=FALSE,  comment=NULL, warning=FALSE, message=FALSE}
p <- qplot(price,data = df, geom = "bar")
p <- p + ggtitle("Price Distribution: AVG =$211")
p <- p + theme(panel.border = element_rect(linetype = "solid", color = "black", fill = NA))
p <- p + theme(plot.title = element_text(colour="grey20",size=20,face="bold"))
p <- p + theme(axis.text = element_text(face="bold"))
p <- p + theme(axis.title.y = element_text(colour="grey20",size=20,face="bold"),
               axis.text.x = element_text(colour="grey20",size=20,face="bold"),
               axis.text.y = element_text(colour="grey20",size=20,face="bold"),  
               axis.title.x = element_text(colour="grey20",size=20,face="bold"))
p <- p + scale_x_continuous( # Update x axis
  name = "Price Distribution",     # Update name, i.e., axis title
  labels = dollar)         # Label axis as dollars. Needs library(scales)
p



p <- qplot(neighbourhood_group, price, data = df, geom = "boxplot", xlab = "Island", ylab = "Price", log = "y")
p <- p + ggtitle("Avg price per Island")
p <- p + theme(plot.title = element_text(colour="grey20",size=20,face="bold"))
p <- p + theme(axis.text = element_text(face="bold"))
p <- p + theme(axis.title.y = element_text(colour="grey20",size=20,face="bold"),
               axis.text.x = element_text(colour="grey20",size=20,face="bold"),
               axis.text.y = element_text(colour="grey20",size=20,face="bold"),  
               axis.title.x = element_text(colour="grey20",size=20,face="bold"))
p

df$price <- factor(df$price)
df$price <- as.numeric(df$price)
median(df$price)
mean(df$price)




```

## Great Hosts per Island

```{r, echo=FALSE,  comment=NULL, warning=FALSE, message=FALSE}
####################################################################################################################
# Create a Table showing Great Hosts per Island
####################################################################################################################
table(df$great_host)


###################################################################################################################
#Plot Great Hosts
###################################################################################################################
#df$great_host

p <- qplot(great_host, data = df, geom = "bar",
           fill = I("yellow"),                     # Add a light blue fill
           color = I("red"),                       # Make color of rectangles red
           alpha = I(0.5))
p <- p + ggtitle("Total Great Hosts")
p <- p + theme(panel.border = element_rect(linetype = "solid", color = "black", fill = NA))
p <- p + theme(plot.title = element_text(colour="grey20",size=20,face="bold"))
p <- p + theme(axis.text = element_text(face="bold"))
p <- p + theme(axis.title.y = element_text(colour="grey20",size=20,face="bold"),
               axis.text.x = element_text(colour="grey20",size=20,face="bold"),
               axis.text.y = element_text(colour="grey20",size=20,face="bold"),  
               axis.title.x = element_text(colour="grey20",size=20,face="bold"))
p <- p + scale_x_discrete( # Update x axis
  name = "")     # Update name, i.e., axis title
p

#p <- qplot(longitude, latitude, data = df, color = neighbourhood_group, na.rm = TRUE, xlab="Longitude", ylab="Latitude") 
#p <- p + ggtitle("Where are the great hosts per island 6975 / 22725")
#p 

p <- qplot(great_host, data = df, geom = "bar", facets = . ~ neighbourhood_group,
           fill = I("yellow"),                     # Add a light blue fill
           color = I("red"),                          # Make color of rectangles red
           alpha = I(0.5))                            # Set 50% transparancy    
p <- p + ggtitle("Distribution of great Hosts per island")
p <- p + theme(panel.border = element_rect(linetype = "solid", color = "black", fill = NA))
p <- p + theme(plot.title = element_text(colour="grey20",size=20,face="bold"))
p <- p + theme(axis.text = element_text(face="bold"))
p <- p + theme(axis.title.y = element_text(colour="grey20",size=20,face="bold"),
               axis.text.x = element_text(colour="grey20",size=20,face="bold"),
               axis.text.y = element_text(colour="grey20",size=20,face="bold"),  
               axis.title.x = element_text(colour="grey20",size=20,face="bold"))
p <- p + scale_x_discrete( # Update x axis
  name = "")     # Update name, i.e., axis title
p <- p + theme(axis.text.x = element_text(angle = 90))
p

#qplot(price, data = df, geom = "density", log = "x", facets = neighbourhood_group ~ great_host)
####################################################################################################################

```

## Do better Review Scores mean more expensive

```{r, echo=FALSE,  comment=NULL, warning=FALSE, message=FALSE}

####################################################################################################################
#Review Scores
####################################################################################################################
rs <- subset(df, reviewscores >= 95)
p <- qplot(reviewscores, data=rs, geom="bar", facets = . ~ neighbourhood_group, xlab = "Review Scores")
p <- p + ggtitle("Review Scores per Island: 95% or greater")
p <- p + theme(panel.border = element_rect(linetype = "solid", color = "black", fill = NA))
p <- p + theme(plot.title = element_text(colour="grey20",size=20,face="bold"))
p <- p + theme(axis.text = element_text(face="bold"))
p <- p + theme(axis.title.y = element_text(colour="grey20",size=20,face="bold"),
               axis.text.x = element_text(colour="grey20",size=20,face="bold"),
               axis.text.y = element_text(colour="grey20",size=20,face="bold"),  
               axis.title.x = element_text(colour="grey20",size=20,face="bold"))
p


p <- qplot(reviewscores, price, data=rs, geom="jitter", facets = . ~ neighbourhood_group, xlab = "Review Scores")
p <- p + ggtitle("Review Scores per Island: 95% or greater")
p <- p + theme(panel.border = element_rect(linetype = "solid", color = "black", fill = NA))
p <- p + theme(plot.title = element_text(colour="grey20",size=20,face="bold"))
p <- p + theme(axis.text = element_text(face="bold"))
p <- p + theme(axis.title.y = element_text(colour="grey20",size=20,face="bold"),
               axis.text.x = element_text(colour="grey20",size=20,face="bold"),
               axis.text.y = element_text(colour="grey20",size=20,face="bold"),  
               axis.title.x = element_text(colour="grey20",size=20,face="bold"))
p
####################################################################################################################

```




## Question 1: Does it get more expensive closer to shore
## It appears yes

```{r, echo=FALSE,  comment=NULL, warning=FALSE, message=FALSE}



qplot(longitude, price, data = df, geom = "point")
qplot(price, data = df, geom ="density")
qplot(longitude, price, data = df, geom = "bin2d", log = "y")
qplot(longitude, price, data = df, geom = "density2d", log = "y")

```



## Let's filter for Hawaii and great hosts
## 

```{r, echo=FALSE,  comment=NULL, warning=FALSE, message=FALSE}

df <- subset(df, neighbourhood_group == "Hawaii" & great_host == "Great Host")
p <- qplot(longitude, latitude, data = df, color = neighbourhood_group, na.rm = TRUE, xlab="Longitude", ylab="Latitude") 
p <- p + scale_color_hue(name = "Island")
p <- p + ggtitle("Overview of AirBNB for Hawaii and Great Host")
p <- p + theme(panel.border = element_rect(linetype = "solid", color = "black", fill = NA))
p <- p + geom_text(aes(label=ifelse(latitude<(19.01),as.character(neighbourhood_group),'')),hjust=-.1,vjust=-7, color = "blue", size= 8)
p <- p + geom_text(aes(label=ifelse(latitude>(21.7) & latitude<(21.8),as.character(neighbourhood_group),'')),hjust=-.5,vjust=1, color = "blue", size = 8)
p <- p + geom_text(aes(label=ifelse(latitude>(21.86) & latitude<(21.87),as.character(neighbourhood_group),'')),hjust=-.2,vjust=1, color = "blue", size = 8)
p <- p + geom_text(aes(label=ifelse(latitude>(21.13) & latitude<(21.14),as.character(neighbourhood_group),'')),hjust=-.5,vjust=4, color = "blue", size = 8)
p <- p + theme(plot.title = element_text(colour="grey20",size=20,face="bold"))
p <- p + theme(axis.text = element_text(face="bold"))
p <- p + theme(axis.title.y = element_text(colour="grey20",size=20,face="bold"),
               axis.text.x = element_text(colour="grey20",size=20,face="bold"),
               axis.text.y = element_text(colour="grey20",size=20,face="bold"),  
               axis.title.x = element_text(colour="grey20",size=20,face="bold"))
p

```


## Does it get more expensive closer to shore for Hawaii
## It appears yes

```{r, echo=FALSE,  comment=NULL, warning=FALSE, message=FALSE}



qplot(longitude, price, data = df, geom = "point")
qplot(price, data = df, geom ="density")
qplot(longitude, price, data = df, geom = "bin2d", log = "y")
qplot(longitude, price, data = df, geom = "density2d", log = "y")

```






















